Leader functions and qualities across 59 cultures: an exploratory analysis

### 2019-12-01 leadership_across_cultures_contexts.Rmd
Corrected spelling

Within each cluster analyses, we computed the distance between two variables as $1 - cor (v_i, v_j)$, so that the distance between highly correlated variables is close to $0$, the distance between uncorrelated variables is close to $1$, and the distance between uncorrelated variables us close to $2$. Using the pvclust package [@suzuki_pvclust2006] we clustered the variables using the matrix of all pairwise distances using the Ward agglomeration algorithm.

@@

The leader quality cluster analysis (Figure \@ref(fig:cluster)A) identified two top-level clusters. Within each top-level cluster we identify two interpretable and generally strongly supported sub-clusters which are annotated with text. Within the first top-level cluster, we identify a dominant/shaman cluster, which suggests a strong link between dominance-based leadership (i.e., *Authority*, *Aggressiveness*, *Feared*) and shamanism (i.e., *Supernatural*, *Artistic performance*). The second sub-cluster includes qualities of leaders related to material, social, and reproductive success (i.e., *Wealthy*, *Social contacts*, and *Polygnous*). Within the second top-level, we identify a sub-cluster of positive personal characteristics (i.e., *Knowledgeable/intelligent*, *Proper behavior*, *Charisma*, *Fairness*) and a sub-cluster of measures related to social strategies and health (i.e., *Strategic planner*, *Generosity*, *Physically formidable*).

@@

A logistic PCA on the `r length(quality_vars)` coded leader quality variables, with rank $k = `r dim(logpca_model_qualities$U)[2]`$ and  $m = `r logpca_model_qualities$m`$ optimized by cross-validation, explained `r signif(logpca_model_qualities$prop_deviance_expl, 3)*100`% of the deviance. We draw on the first two leader quality components for interpretation and additional analyses (see Figure \@ref(fig:qualitiesPCplot)). We identify the first component (PC1) as a *Prestige vs. Dominance* measure. Low values on Leader Qualities PC 1 are strongly associated with the variables *Experienced/accomplished* and *Knowledgeable/intelligence* and moderately associated with *Confidence* and *Bravery*. High values are strongly associated with *Polygnous* and *Authority* and moderately associated with *High quality spouse*. Therefore, the "dominance" side of the *Prestige vs. Dominance* component is also closely liked to reproduction, as demonstrated in the cluster analysis (Figure \@ref(fig:cluster)). Measures of physical formidability and aggressiveness are also moderately associated with the "dominance" side of this component. The second leader qualities component we identify as an *In-group favoritism vs. Social Status* measure. Low values on Leader Qualities PC 2 are strongly associated with *Aggressiveness*, *Bravery*, *Xenophobia*, and *Loyalty*. High Values are strongly associated with *High status* and moderately associated with *Fairness*, *Age*, and *Decisiveness/decision-making*. Therefore, Leader Qualities PC 2 also reflects elements of a prestige-dominance distinction in contrasting aggressiveness with high social status.

>>>>

Within each cluster analyses, we computed the distance between two variables as $1 - cor (v_i, v_j)$, so that the distance between highly correlated variables is close to $0$, the distance between uncorrelated variables is close to $1$, and the distance between negatively correlated variables is close to $2$. Using the pvclust package [@suzuki_pvclust2006] we clustered the variables using the matrix of all pairwise distances using the Ward agglomeration algorithm.

@@

The leader quality cluster analysis (Figure \@ref(fig:cluster)A) identified two top-level clusters. Within each top-level cluster we identify two interpretable and generally strongly supported sub-clusters which are annotated with text. Within the first top-level cluster, we identify a dominant/shaman cluster, which suggests a strong link between dominance-based leadership (i.e., *Authority*, *Aggressiveness*, *Feared*) and shamanism (i.e., *Supernatural*, *Artistic performance*). The second sub-cluster includes qualities of leaders related to material, social, and reproductive success (i.e., *Wealthy*, *Social contacts*, and *Polygynous*). Within the second top-level, we identify a sub-cluster of positive personal characteristics (i.e., *Knowledgeable/intelligent*, *Proper behavior*, *Charisma*, *Fairness*) and a sub-cluster of measures related to social strategies and health (i.e., *Strategic planner*, *Generosity*, *Physically formidable*).

@@

A logistic PCA on the `r length(quality_vars)` coded leader quality variables, with rank $k = `r dim(logpca_model_qualities$U)[2]`$ and  $m = `r logpca_model_qualities$m`$ optimized by cross-validation, explained `r signif(logpca_model_qualities$prop_deviance_expl, 3)*100`% of the deviance. We draw on the first two leader quality components for interpretation and additional analyses (see Figure \@ref(fig:qualitiesPCplot)). We identify the first component (PC1) as a *Prestige vs. Dominance* measure. Low values on Leader Qualities PC 1 are strongly associated with the variables *Experienced/accomplished* and *Knowledgeable/intelligence* and moderately associated with *Confidence* and *Bravery*. High values are strongly associated with *Polygynous* and *Authority* and moderately associated with *High quality spouse*. Therefore, the "dominance" side of the *Prestige vs. Dominance* component is also closely liked to reproduction, as demonstrated in the cluster analysis (Figure \@ref(fig:cluster)). Measures of physical formidability and aggressiveness are also moderately associated with the "dominance" side of this component. The second leader qualities component we identify as an *In-group favoritism vs. Social Status* measure. Low values on Leader Qualities PC 2 are strongly associated with *Aggressiveness*, *Bravery*, *Xenophobia*, and *Loyalty*. High Values are strongly associated with *High status* and moderately associated with *Fairness*, *Age*, and *Decisiveness/decision-making*. Therefore, Leader Qualities PC 2 also reflects elements of a prestige-dominance distinction in contrasting aggressiveness with high social status.
